### 2021-12-02 input/proposal.tex
about algo

\TODO{Rework this to remove repetitions. Since implem is close from
  section.}

Algorithm~\ref{algo:ascast} shows the instructions implementing
dynamic consistent partitioning. To implement staleness, each \process
maintains a vector of versions that associates the respective known
local counter of each known source, or has-been source. Its size
increases linearly with the number of sources that the \process has
ever known, which is the number of \processes in the system
$\mathcal{O}(V)$ in the worst case.  Nevertheless, following the
principles of scoped broadcast, we expect that every \process only
acknowledges a small subset of sources.  In addition, each message
carries the identifier and counter of each \node that forwarded
it. Again, following the principles of scoped broadcast, we expect
that versioned paths remain small. Using these data structures, each
\process delivers and forwards only new messages that improve their
best known partition (see Line~\ref{line:ascast_improve}).

This vector of versions constitutes a summary of known progress of
other \nodes. It discards stale messages and it enables detecting
possible blocking of staleness notifications: the current best
partition comes from a neighbor that just delivered a stale message
(see Line~\ref{line:ascast_detect}). Figure~\ref{fig:problem} shows
that such a behavior may hide a delete notification. The detecting
\process must act to reflect this possible change: all \processes
whose last message comes from the detecting \process must remove it
and accept the competition of other \processes. Since this behavior is
similar to that of source deletion, we implement it in a similar way.

A \process that removes its self-appointed status of source, or
detects a possible inconsistency, increments its local counter and
disseminates a delete notification message $\delta_{s, c}$
piggybacking its identifier $s$ and counter $c$. Any \process that
receives such delete notification delivers and forwards it if its best
message contains the has-been source or detector associated to a lower
counter (see Line~\ref{line:ascast_delete}). This allows $\delta$
messages to quickly reach every downstream \process.

As stated in Theorem~\ref{theo:dcp}, dynamic consistent partitioning
not only requires eventual purging of stale messages, but also the
retrieval of the best up-to-date partitions.  In that regard, $\delta$
messages have a dual use: since they already reach the borders of
their partition by design, they trigger a competition when reaching
bordering.  When a \process receives a delete notification from a
neighbor $q$ but does not deliver it, it assumes that the just
received $\delta$ message purged possibly stale messages, creating a
gap that its own best partition could help fill. Consequently, it
sends to $q$ its own best $\alpha$ message (see
Line~\ref{line:ascast_echo}).

@@

Algorithm~\ref{algo:edges} enables dynamic consistent partitioning in
dynamic networks where \processes can join, leave, or crash at any
time. Removing a \process is equivalent to removing all its edges, and
adding a \process is equivalent to add as many edges as necessary.
Adding an edge between two \processes may only improve the shortest
path of one of these \processes. Therefore, it triggers a competition
between the two \processes. If one or the other \process improves, the
propagation of $\alpha$ messages operates normally.  Removing an edge
between two \processes may invalidate the shortest path of one of
involved \processes plus downstream \processes. As a side effect,
removing an edge may also impair the propagation of a partition
delete. To implement this behavior, \NAME uses its implementation of
$\delta$ messages. This enables \NAME even in dynamic networks subject
to physical partitioning.

%% \NAME guarantees dynamic consistent partitioning by making extensive
%% use of \process-to-\process communications. It requires
%% \begin{inparaenum}[(i)]
%% \item purging stale control information, hence propagating $\delta$
%%   messages; and
%% \item retrieving the closest source, hence propagating $\alpha$
%%   messages.
%% \end{inparaenum}
In terms of number of messages, in the average case, a \process $i$
chosen uniformly at random among all \processes creates a logical
partition. Its messages $\alpha_i$ spread through the network until
reaching \processes that belong to another partition closer to
them. This splits partitions in half on average. Overall, the $a^{th}$
new partition comprises \smash{$\mathcal{O}(\frac{|V|}{2^{\lfloor
      \log_2 a \rfloor}})$} \processes. This decreases every new
partition until reaching one \process per new partition: its
source. The average number of messages per \process is
\smash{$\mathcal{O}(\frac{\overline{|O|}}{2^{\lfloor \log_2 a
      \rfloor}})$}, where \smash{$\overline{|O|}$} is the average
number of neighbors per \process.
% \TODO{Multiple receipt and
% multiple
%  delivery imply more messages (receipt bounded by $|O|$ as well).}
Deleting the $a^{th}$ partition generates twice as many messages as
creating the $a^{th}$ partition: $\delta$ messages travel through the
partition, then $\alpha$ messages compete to fill the gap.  Overall,
the communication complexity shows that \NAME scales well to the
number of partitions.
%% In the
%% worst-case, every new partition includes all but one \process
%% belonging to the previous partition. The total number of messages
%% after the $a^{th}$ new partition is $\mathcal{O}(\overline{|O|}\cdot
%% a^2)$. As for the average-case, the number of messages for the
%% partition deletion is identical to the number of messages of the
%% corresponding partition creation.
%% \begin{algorithm}[h!]
%%   \input{input/algoedges.tex}
%%   \caption{\label{algo:edges}\NAME at \Process~$p$ in dynamic networks.}
%% \end{algorithm}

>>>>

Algorithm~\ref{algo:ascast} shows the instructions of our
implementation called \NAME that enables dynamic consistent
partitioning. It implements forwarding of best (see
Line~\ref{line:fb}) and echos (see Line~\ref{line:echo}). To implement
forwarding of staleness downstream as stated in
Definition~\ref{def:fsplus},
\begin{inparaenum}[(A)]
\item each \process maintains a vector of versions that associates the
  respective known local counter of each known source, or has-been
  source. It constitutes a summary of known progress of other
  \processes; 
\item each notification message $\alpha$ carries the list of
  identifier and counter of each \node that forwarded it. In the worst
  case, both these structures grow linearly with the number of
  \processes in the system $\mathcal{O}(V)$. Nevertheless, following
  the principles of scoped broadcast, we expect that \processes only
  acknowledge a small subset of sources and messages;
\item each staleness notifications $\delta$ only carry the identifier
  and counter of the \process~--~source or detector~--~that generated
  it. Only downstream \processes may deliver such message, since they
  carry the identifier of the generator. 
\end{inparaenum}
To implement the detection as stated in Lemma~\ref{lem:detector}, each
\process only requires to know the direct parent of its best delivered
message which is already included in the piggybacked path of this
message. Receiving a message known to be stale from this parent
triggers the generation of staleness notifications that force
downstream \processes to reevaluate the partition they are in (see
Line~\ref{line:detect}).

@@

By reusing \NAME's default behavior, Algorithm~\ref{algo:edges}
enables dynamic consistent partitioning in dynamic networks where
\processes can join, leave, or crash at any time.  Adding a \process
is equivalent to add as many edges as necessary. Since adding an edge
may improve shortest paths, the protocol triggers a competition using
echos of Definition~\ref{def:fbplus}.  Removing a \process is
equivalent to removing all its edges. Removing an edge between two
\processes may invalidate the shortest path of one of involved
\processes plus downstream \processes, or impair the propagation of
staleness notifications. Therefore, the protocol reuses its detection
mechanism of Lemma~\ref{lem:detector} as a solution.
Algorithm~\ref{algo:edges} even extends \NAME to dynamic networks
subject to physical partitioning.

%% \NAME guarantees dynamic consistent partitioning by making extensive
%% use of \process-to-\process communications. It requires
%% \begin{inparaenum}[(i)]
%% \item purging stale control information, hence propagating $\delta$
%%   messages; and
%% \item retrieving the closest source, hence propagating $\alpha$
%%   messages.
%% \end{inparaenum}
\TODO{TODOUBLECHECK}. In terms of number of messages, in the average
case, a \process $i$ chosen uniformly at random among all \processes
creates a logical partition. Its messages $\alpha_i$ spread through
the network until reaching \processes that belong to another partition
closer to them. This splits partitions in half on average. Overall,
the $a^{th}$ new partition comprises
\smash{$\mathcal{O}(\frac{|V|}{2^{\lfloor \log_2 a \rfloor}})$}
\processes. This decreases every new partition until reaching one
\process per new partition: its source. The average number of messages
per \process is \smash{$\mathcal{O}(\frac{\overline{|O|}}{2^{\lfloor
      \log_2 a \rfloor}})$}, where \smash{$\overline{|O|}$} is the
average number of neighbors per \process.
% \TODO{Multiple receipt and
% multiple
%  delivery imply more messages (receipt bounded by $|O|$ as well).}
Deleting the $a^{th}$ partition generates twice as many messages as
creating the $a^{th}$ partition: $\delta$ messages travel through the
partition, then $\alpha$ messages compete to fill the gap.  Overall,
the communication complexity shows that \NAME scales well to the
number of partitions.
%% In the
%% worst-case, every new partition includes all but one \process
%% belonging to the previous partition. The total number of messages
%% after the $a^{th}$ new partition is $\mathcal{O}(\overline{|O|}\cdot
%% a^2)$. As for the average-case, the number of messages for the
%% partition deletion is identical to the number of messages of the
%% corresponding partition creation.
%% \begin{algorithm}[h!]
%%   \input{input/algoedges.tex}
%%   \caption{\label{algo:edges}\NAME at \Process~$p$ in dynamic networks.}
%% \end{algorithm}
